---
title: "<h1 style = 'color: grey;'> Constrained Optimal Sample Allocation Tutorial </h1>"
author: "Metin Bulus"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{COSA Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cosa)
```

The following example demonstrates how to perform constrained optimal sample allocation (COSA) for a two-level cluster randomized trial (CRT) and for a corresponding cluster-level regression discontinuity (CRD) design. Number of clusters is calculated where treatment and control units are associated with differential marginal costs. This result in a cost-efficient treatment group allocation (`p`) for the CRT along with the optimal number of clusters. However, in a comparable CRD design proportion of treatment group allocation cannot be optimized as the cutoff and the scoring variable is predetermined. 

*Cluster Randomized Trial (CRT)*

```{r, message=FALSE}
# cost constrained - optimize p and n2
# Specifying rhots as 0 allows COSA in CRT
crt <- cosa.crd2r2(rhots = 0,
                   constrain = "cost", cost = 15000,
                   cn1 = c(5, 2), cn2 = c(50, 20),
                   es = .20, rho2 = .20,
                   g2 = 5, r21 = .20, r22 = .30,
                   p = NULL, n1 = 24,  n2 = NULL)
```

*Cluster-level Regression Discontinuity (CRD) Study*

```{r, message=FALSE}
# cost constrained - optimize n2
crd <- cosa.crd2r2(constrain = "cost", cost = 15000,
                   cn1 = c(5, 2), cn2 = c(50, 20),
                   es = .20, rho2 = .20,
                   g2 = 5, r21 = .20, r22 = .30,
                   p = .387, n1 = 24,  n2 = NULL)
```

*Comparison via MDES and Power Plots*
```{r, message=FALSE, fig.width=7, fig.height=5}
par(mfrow = c(1, 2), mai = c(1, .9, .5, .2))

plot(crt, ypar = "mdes", xpar = "n2",
     ylim = c(0, 1), xlim = c(10, 200), 
     ylab = "MDES (Power = .80)", xlab = "Number of Clusters",
     main = "CRT", locate = TRUE)
plot(crd, ypar = "mdes", xpar = "n2",
     ylim = c(0, 1), xlim = c(10, 200),
     ylab = "MDES (Power = .80)", xlab = "Number of Clusters",
     main = "CRD", locate = TRUE)

plot(crt, ypar = "power", xpar = "n2",
     ylim = c(0, 1), xlim = c(10, 200), 
     ylab = "Power (ES = .20)", xlab = "Number of Clusters",
     main = "CRT", locate = TRUE)
plot(crd, ypar = "power", xpar = "n2",
     ylim = c(0, 1), xlim = c(10, 200), 
     ylab = "Power (ES = .20)", xlab = "Number of Clusters",
     main = "CRD", locate = TRUE)

```

--o--
