---
title: "Constrained Optimal Sample Allocation in Multilevel Randomized Trials"
author: "Metin Bulus"
date: "February 5, 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cosa Vignettes}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cosa)
```

Specify `rhots = 0` for constrained optimal sample allocation in multilevel randomized trials. Of particular interest is the optimization of proportion of treatment allocation (`p`) under unequal marginal cost assumption. In the following examples `p` is optimized along with the sample size at the randomization level. 

# Treatment at level 1

## Block/Multisite Two-level Random Assignment (BIRA2r1) Design
```{r, message=FALSE, warning=FALSE}
# cost constrained - optimize n1 and p
cosa.bird2r1(constrain = "cost", cost = 2000, rhots = 0, 
             cn1 = c(5, 2), cn2 = 10,
             es = .20, rho2 = .20, omega2 = .20,
             r21 = .20, r2t2 = .20, g2 = 1, 
             n2 = 150)
```

## Block/Multisite Three-level Random Assignment (BIRA3r1) Design
```{r, message=FALSE}
# cost constrained - optimize n1 and p
cosa.bird3r1(constrain = "cost", cost = 10000, rhots = 0, 
             cn1 = c(5, 3), cn2 = 10, cn3 = 30,
             es = .20, rho2 = .20, rho3 = .10, omega2 = .20, omega3 = .10,
             r21 = .20, r2t2 = .20, r2t3 = .20, g3 = 1, n2 = 3, n3 = 50)
```

## Block/Multisite Four-level Random Assignment (BIRA4r1) Design
```{r, message=FALSE}
# cost constrained - optimize n1 and p
cosa.bird4r1(constrain = "cost", cost = 50000, rhots = 0, 
             cn1 = c(5, 3), cn2 = 10, cn3 = 30, cn4 = 50,
             es = .20, rho2 = .20, rho3 = .10, rho4 = .05,
             omega2 = .20, omega3 = .10, omega4 = .10,
             r21 = .20, r2t2 = .20, r2t3 = .20, r2t4 = .20,
             g4 = 1, n2 = 3, n3 = 20, n4 = 5)

```

# Treatment at level 2

## Cluster-level Random Assignment (CRA2r2) Design
```{r, message=FALSE}
# cost constrained - optimize n2 and p
cosa.crd2r2(constrain = "cost", cost = 10000, rhots = 0, 
            cn1 = 5, cn2 = c(10, 8),
            es = .20, rho2 = .20, r21 = .20, r22 = .30,
            g2 = 1, n1 = 10)
```

## Block/Multisite Three-level Random Assignment (BCRA3r2) Design
```{r, message=FALSE}
# cost constrained - optimize n2 and p
cosa.bcrd3r2(constrain = "cost", cost = 10000, rhots = 0, 
             cn1 = c(5, 2), cn2 = 10, cn3 = 30,
             es = .20, rho2 = .20, rho3 = .10, omega3 = .10,
             r21 = .20, r22 = .20, r2t3 = .20,
             g3 = 1, n1 = 25, n3 = 20)
```

## Block/Multisite Four-level Random Assignment (BCRA4r2) Design
```{r, message=FALSE}
# cost constrained - optimize n2 and p
cosa.bcrd4r2(constrain = "cost", cost = 100000, rhots = 0, 
             cn1 = c(5, 4), cn2 = 10, cn3 = 30, cn4 = 50,
             es = .20, rho2 = .20, rho3 = .10, rho4 = .05,
             omega3 = .10, omega4 = .20,
             r21 = .20, r22 = .20, r2t3 = .20, r2t4 = .20,
             g4 = 1, n1 = 25, n3 = 50, n4 = 5)
```

# Treatment at level 3

## Cluster-level Random Assignment (CRA3r3) Design
```{r, message=FALSE}
# cost constrained - optimize n3 and p
cosa.crd3r3(constrain = "cost", cost = 50000, rhots = 0, 
            cn1 = 5, cn2 = 10, cn3 = c(50, 20),
            es = .20, rho2 = .20, rho3 = .10,
            r21 = .20, r22 = .30, r23 = .40,
            g3 = 1, n1 = 25, n2 = 3)
```

## Block/Multisite Four-level Random Assignment (BCRA4r3) Design
```{r, message=FALSE}
# cost constrained - optimize n3 and p
cosa.bcrd4r3(constrain = "cost", cost = 50000, rhots = 0, 
             cn1 = c(5, 3), cn2 = 10, cn3 = 30, cn4 = 50,
             es = .20, rho2 = .20, rho3 = .10, rho4 = .05, omega4 = .20,
             r21 = .20, r22 = .20, r23 = .20, r2t4 = .20,
             g4 = 1, n1 = 25, n2 = 3, n4 = 5)
```

# Treatment at level 4

## Cluster-level Random Assignment (CRA4r4) Design
```{r, message=FALSE}
# cost constrained - optimize n4 and p
cosa.crd4r4(constrain = "cost", cost = 500000, rhots = 0, 
            cn1 = 5, cn2 = c(10, 5), cn3 = c(30, 10), cn4 = c(50,30),
            es = .20, rho2 = .20, rho3 = .10, rho4 = .05,
            r21 = .20, r22 = .30, r23 = .40, r24 = .50,
            g4 = 1, n1 = 25, n2 = 3, n3 = 10)
```
